Fast Algorithms for Intersection of Nonmatching Grids Using \plucker coordinates.

### 2016-09-30 paper_intersections.tex
intersection 13d fished

\title{Intersection of non-matching grids of different dimensions}
% Self-intersection of non-matching grid of mixed dimensions

@@

\begin{abstract}
TODO: An abstract.
\end{abstract}

@@

\subsection{Intersection Line-Triangle (1D-2D)}
Let us consider a line $p$ with parametric equation $\vc X = \vc A + t\vc u$, $t\in\R$ and a triangle $T$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. 
\begin{lemma}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $X$ on the $p$ and inside the triangle $T$. 
The barycentric coordinates of $X$ on $T$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ have \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
The point $X$ is inside of $T$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $X$ on $T$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

@@

\subsection{Intersection Line-Teraherdon (1D-3D)}
\noteJB{Has to be completed yet, however we consider that we obtain 0,1, or 2 distinguish ICs for all cases including various touch cases. Anyway every IC results from 
some non-degenerate 1d-2d intersection. We also assume that the ICs are in order according to the orientation of the side which is
counterclockwise aroung the triangle.}
We consider a line $p$ and a tetraherdon...
Tetrahedron has six edges, so six \plucker coordinates and inner products are computed at most. These data are
passed to the 1D-2D algorithm described above, which is applied to the every face of the tetrahedron face.

There might be two intersection corners maximally, which are finally checked that they belong o abscissas.
If the line has its boundary point inside the tetrahedron, the intersection line must be cut
and coordinates of intersection corner are interpolated.
Pathologic cases are again carefully processed so we obtain the most extensive topology information we can.
See \cite{fris_dp_2015} for detailed description of all possible cases.

>>>>

%\title{Algorithm for Intersections of Nonmatching Grids of Different Dimensions}
\title{Fast Algorithms for Intersection of Nonmatching Grids Using \plucker coordinates.}

@@

\begin{abstract}
\end{abstract}

@@

\subsection{Intersection Line-Triangle (1D-2D)}
Let us consider a line segment $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,\ t\in (0,1)
\end{equation}
and a triangle $T$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. 
\begin{lemma}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $X$ on the $p$ and inside the triangle $T$. 
The barycentric coordinates of $X$ on $T$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ have \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
The point $X$ is inside of $T$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $X$ on $T$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

@@

\subsection{Intersection Line-Teraherdon (1D-3D)}
In this section we consider intersection of a line segment $p$ given by the parametric equation \eqref{eq:line_parametric} with a tetrahedron 
$S_3$. The used algorithm is based on 1d-2d algorithm and closely follows \cite{platis_fast_2003}. Our modification takes into account
intersection with a line segment and consistently propagates topological position of ICs.

\begin{algorithm}
  \caption{1d-3d intersection}
  \label{algo:13d}

  \DontPrintSemicolon
  \SetKw{and}{and}
  \SetKw{continue}{continue}
  \SetKw{break}{break}
  \KwIn{Tetrahedron $S_3$, line segment $p$.}
  \KwOut{List of ICs on sorted along $p$.}  
  $I=\{\}$ 
  \For{unmarked face $f$ of $S_3$}{
    $L$ = intersection($p,\ f$)\;
    \lIf{ $L$ is none or degenerate}{\continue}
    \lIf{ $L$ is inside the edge $e$}{
        set $\tau_{S_4} = (1,e)$\;
        mark faces coincident with $e$
    }
    \ElseIf{ $L$ is at the vertex $v$}{
        set $\tau_{S_4} = (0,v)$\;
        mark faces coincident with $v$
    }
    append $L$ to $I$
    \lIf{$\abs{I}=2$}{\break} 
  }
  \lIf{$\abs{I}=1$ \and   $I$ is outside of $p$}{ erase $I$}
  \ElseIf{$\abs{I}=2$}{    
    trim intersection with respect to the line segment $p$ \label{line:trimming}
  }
\end{algorithm}
Algorithm \ref{algo:13d} first compute line-face intersections for every face of $S_4$ avoiding duplicate computation of ICs on edges and at vertices and
skipping remaining faces once two ICs are found. Tetrahedron has six edges, so 7 \plucker coordinates and 6 inner products are computed at most. Precomputed 
coordinates and products are passed the 1D-2D algorithm which is performed fro the whole line $p$. After collecting line-tetrahedron ICs,
we do the line segment trimming \ref{line:trimming}. If both ICs are out of the line segment $p$ we eliminate both of them. If one of the ICs if out of $p$
we use the closest end point of the line segment instead and interpolate barycentric coordinates of the IC on $S_4$. The topological positions are updated as well.
The result of the algorithm are zero up to two ICs sorted by the parameter $t$ of the line $p$.
